---
title: "EEG_Permutation_Analyses"
author: "Daniele Scanzi"
date: "14/11/2021"
output: html_document
---

```{r include=FALSE}
### Set up the environment  ###

# Load all the necessary packages
library(here)
library(tidyverse)
library(janitor)
library(ggplot2)
library(Rmisc)
library(RColorBrewer)
library(ggthemes)
library(permuco)
library(brms)
library(ggmcmc)
library(mcmc)
library(ggthemr)

# Plotting Theme
palette <- define_palette(
  swatch = c('#111111', '#E84646', '#233B43', '#C29365', '#168E7F', '#65ADC2', '#109B37', '#DB784D', '#f1c40f'),
  gradient = c(lower = '#E84646', upper = '#109B37'))
ggthemr(palette, layout = 'scientific')

# Distractors sides
right.distractors <- c(2,3,4,5,6,7,8)
left.distractors <- c(10,11,12,13,14,15,16)
```

```{r}
### LOAD AND PREPARE DATA ###
# Data was extracted through MATLAB and required some data wrangling


# HIGH WMC DATA
# Dataset should be contained in a folder called Datasets within the current directory
highWMCdata <- read_csv(here::here('Datasets', 'EEG_SingleTrial_High.csv')) %>% 
  # Pivot longer 
  tidyr::pivot_longer(cols      = '-200':dplyr::last_col(),
                      names_to  = 'time_point',
                      values_to = 'signal') %>% 
  # Clean column names (all lower case, style:name1_name2)
  janitor::clean_names() %>% 
  # Select relevant coloumns
  dplyr::select(trial,part_column, wmc_column, electrode, electrode_position, distractor_type, distractor_colour, distractor_position, time_point, signal) %>% 
  dplyr::rename(wmc = wmc_column,
                id  = part_column) %>% 
  dplyr::mutate(wmc = replace(wmc, wmc == 'hig', 'high'))


# LOW WMC DATA
# Dataset should be contained in a folder called Datasets within the current directory
lowWMCdata <- read_csv(here::here('SingleTrials_Datasets', 'Low_SingleTrial_All.csv')) %>% 
  # Pivot longer
  tidyr::pivot_longer(cols      = '-200':dplyr::last_col(),
                      names_to  = 'time_point',
                      values_to = 'signal') %>% 
  # Clean column names
  janitor::clean_names() %>% 
  dplyr::select(trial,part_column, wmc_column, electrode, electrode_position, distractor_type, distractor_colour, distractor_position, time_point, signal) %>% 
  dplyr::rename(wmc = wmc_column,
                id  = part_column) %>% 
  dplyr::mutate(wmc = replace(wmc, wmc == 'hig', 'high'))

# Join the data in a single dataset
trialData <- rbind(highWMCdata,
                   lowWMCdata) %>% 
  # Trasform relevant columns into factors
  dplyr::mutate(id                  = factor(id),
                wmc                 = factor(wmc),
                electrode           = factor(electrode),
                electrode_position  = factor(electrode_position),
                distractor_type     = factor(distractor_type),
                distractor_colour   = factor(distractor_colour),
                distractor_position = as_factor(distractor_position),
                time_point          = as.numeric(time_point))
```

# Compute subtraction

Ipsi - Contralateral in reference to the side of the presentation of the distractor
```{r}
### Compute subtracted waveforms ###
trialSubtracted <- trialData %>% 
  dplyr::select(!electrode) %>% 
  tidyr::pivot_wider(names_from  = electrode_position,
                     values_from = signal) %>% 
  # Compute contra - Ipsi
  dplyr::mutate(subtracted = contra - ipsi) %>% 
  dplyr::group_by()
```

```{r}
### Averages ###

# Compute across trials - within condition for each subject
subjectAvg <- trialSubtracted %>% 
  dplyr::group_by(wmc, id, distractor_type, distractor_colour, time_point) %>% 
  dplyr::summarise(avg_signal = mean(subtracted)) %>% 
  dplyr::ungroup()

#Compute across trials - within condition
conditionAvg <- trialSubtracted %>% 
  dplyr::group_by(wmc, distractor_type, distractor_colour, time_point) %>% 
  dplyr::summarise(avg_signal = mean(subtracted)) %>% 
  dplyr::ungroup()
```

## Butterfly Plots
```{r}
### Plot butterfly plots for each condition ###

# High WMC
subjectAvg %>% 
  dplyr::filter(wmc == 'high') %>% 
  dplyr::group_by(wmc, id, distractor_type, time_point) %>% 
  dplyr::summarise(avg_signal = mean(avg_signal)) %>% 
  ggplot(aes(x      = time_point, 
             y      = avg_signal,
             fill   = id,
             colour = id)) +
  geom_line(size  = .75,
            alpha = .85) +
  # Set axis ticks
  scale_x_continuous(breaks = c(seq(-200, 800, 200))) +
  scale_y_continuous(breaks = c(seq(-5, 5, 0.5))) + 
  coord_cartesian() +
  # Add orizontal line at 0 and red vertical line at onset of stimulus
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0, colour = 'red') +
   # Set labels
  labs(x     = 'Time (ms)',
       y     = 'Amplitude (mV)',
       title = 'Butterfly plots for each condition P58/P96') +
  theme_tufte(base_size = 12) +
  facet_grid(id~distractor_type) +
  theme(legend.position = 'none') 

# Low WMC
subjectAvg %>% 
  dplyr::filter(wmc == 'low') %>% 
  dplyr::group_by(wmc, id, distractor_type, time_point) %>% 
  dplyr::summarise(avg_signal = mean(avg_signal)) %>% 
  ggplot(aes(x      = time_point, 
             y      = avg_signal,
             fill   = id,
             colour = id)) +
  geom_line(size  = .75,
            alpha = .85) +
  # Set axis ticks
  scale_x_continuous(breaks = c(seq(-200, 800, 200))) +
  scale_y_continuous(breaks = c(seq(-5, 5, 0.5))) + 
  coord_cartesian() +
  # Add orizontal line at 0 and red vertical line at onset of stimulus
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0, colour = 'red') +
   # Set labels
  labs(x     = 'Time (ms)',
       y     = 'Amplitude (mV)',
       title = 'Butterfly plots for each condition P58/P96') +
  theme_tufte(base_size = 12) +
  facet_grid(id~distractor_type) +
  theme(legend.position = 'none')
```

```{r}
### ERPs wor WMC divided by colour
conditionAvg %>% 
  dplyr::group_by(wmc, distractor_type, distractor_colour) %>% 
  ggplot(aes(x      = time_point, 
             y      = avg_signal,
             fill   = wmc,
             colour = wmc)) +
  geom_line(size = 1.2) +
  # Define axis ticks
  scale_x_continuous(breaks = c(seq(-200, 800, 100))) +
  scale_y_continuous(breaks = c(seq(-5, 5, 0.2)),
                     labels = scales::comma) + 
  coord_cartesian() +
  # Add Vertical line at 0
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0, 
             colour     = 'red') +
  # Labels
  labs(x     = 'Time (ms)',
       y     = 'Amplitude (\U003BC V)',
       title = "Grand Average ERP across WMC",
       fill  = 'WMC group') +
  # Set Theme
  theme_tufte(base_size = 12) +
  facet_grid(distractor_type ~ distractor_colour)
```

# Permutation Analyses
```{r}
### Set parameters ###
set.seed(2021) # For replicability
nperm <- 1000  # Number of permutations
```

```{r}
### Prepare data to be used in permuco ###

# Create wide dataset with a coloum for each time point
permuDistrData <- subjectAvg %>% 

  # Pivot wider to prepare data as permuco package requires
  dplyr::group_by(wmc, id, distractor_type, distractor_colour) %>% 
  tidyr::pivot_wider(names_from  = time_point,
                     values_from = avg_signal) %>% 
  # Remove datapoints before stimulus onset and after 456ms
  dplyr::select(!('-200':'-4') & !('456':dplyr::last_col())) %>% 
  dplyr::ungroup()

# divide data into "data" and "design" datasets
distrSignal <- permuDistrData %>% 
  dplyr::select('0':last_col())

distrDesign <- permuDistrData %>% 
  dplyr::select(wmc:distractor_colour)
```

```{r}
### Compute permutation Analyses
permDistrTest <- permuco::clusterlm(distrSignal ~ wmc * distractor_type * distractor_colour + Error(id/(distractor_type * distractor_colour)),
                                      data     = distrDesign,
                                      method   = "Rde_kheradPajouh_renaud",
                                      multcomp = 'troendle',
                                      np       = nperm)

plot(permDistrTest)
permDistrTest
```

```{r}
### Plot significant results

# Significant time windows
distractorPermutation <- permDistrTest[["multiple_comparison"]][["distractor_type"]][["troendle"]][["main"]] %>% 
  dplyr::as_tibble() %>% 
  tibble::rowid_to_column(var = 'time') %>% 
  dplyr::mutate(time = (time * 4) - 4) # Shist time so that starts from 0


distrPermuPlot <- distractorPermutation %>% 
  ggplot(aes(x = time,
             y = statistic)) +
  geom_line(size   = .75, 
            colour = palette$swatch[3]) +
  geom_line(data= dplyr::filter(distractor.permutation, pvalue <= .05 & time < 300),
            aes(x  = time, 
                y  = statistic),
            colour = palette$swatch[2],
            size   = 1.2) +
  geom_line(data = dplyr::filter(distractor.permutation, pvalue <= .05 & time > 300),
            aes(x      = time, 
                y      = statistic),
                colour = palette$swatch[2],
            size       = 1.2) +
  geom_vline(xintercept = c(228, 276, 316, 412),
             colour     = palette$swatch[2],
             size       = 0.5) +
  labs(x = 'Time (ms)',
       y = 'Fisher Statistic')

# ERP plot for distractor type effects
distrTypeErp <- trialSubtracted %>% 
  
  # Extract ERPs for distractor type only (average across all other conditions)
  dplyr::group_by(distractor_type, time_point) %>% 
  dplyr::summarise(avg_signal = mean(subtracted)) %>% 
  # Plot results
  ggplot(aes(x      = time_point, 
             y      = avg_signal,
             fill   = distractor_type,
             colour = distractor_type)) +
  geom_line(size = 1.2) +
  annotate('rect',
           xmin    = 228,
           xmax    = 276,
           ymin    = -Inf,
           ymax    = +Inf,
            fill   = palette$swatch[2],
            colour = 'transparent',
            alpha  = .2) +
    annotate('rect',
           xmin    = 316,
           xmax    = 412,
           ymin    = -Inf,
           ymax    = +Inf,
            fill   = palette$swatch[2],
            colour = 'transparent',
            alpha  = .2) +
  # Add Vertical line at 0
  geom_hline(yintercept = 0, 
             colour     = palette$swatch[1]) +
  geom_vline(xintercept = 0, 
             colour     = palette$swatch[1]) +
  annotate(geom  = 'text', 
           x     = 250,
           y     = 1,
           label = 'N2pc') +
  annotate(geom  = 'text', 
           x     = 360,
           y     = 1,
           label = 'Ptc') +
  scale_x_continuous(breaks = seq(-200, 800, 100)) +
  scale_y_continuous(breaks = seq(-1, 1, .5)) +
  # Labels
  labs(x = 'Time (ms)',
       y = 'Amplitude (mV)',
       colour = 'Distractor Type') +
  scale_colour_ggthemr_d()


combined <- cowplot::plot_grid(distrTypeErp,
                   distrPermuPlot,
                   ncol        = 1, 
                   align       = 'vh', 
                   rel_heights = c(1.5,1))

```

### Effects of WMC

The effects of WMC were not significant but visual inspection seemed to tell a different story. Thus they were checked out of precaution. Nonetheless, no significant effects were found.

```{r}
### Plot Effects of WMC (Other condition pooled)
trialSubtracted %>% 
  
  # Extract ERPs for WMC only (average across all other conditions)
  dplyr::group_by(wmc, time_point) %>% 
  dplyr::summarise(avg_signal = mean(subtracted)) %>% 
  # Plot results
  ggplot(aes(x      = time_point, 
             y      = avg_signal,
             fill   = wmc,
             colour = wmc)) +
  geom_line(size = 1.2) +
  # Set colours
  scale_colour_manual(values = mycolours,
                      name   = 'WMC') +
  scale_fill_manual(values = mycolours) +
  # Define axis ticks
  scale_x_continuous(breaks = c(seq(0, 455, 50))) +
  scale_y_continuous(breaks = c(seq(-5, 5, 0.2)),
                     labels = scales::comma) + 
  coord_cartesian(xlim = c(0, 500)) +
  # Add Vertical line at 0
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0, colour = 'red') +
  # Labels
  labs(x = 'Time (ms)',
       y = 'Amplitude (mV)',
       title = "Working Memory Capacity ERPs") +
  scale_colour_ggthemr_d()
```


```{r}
### Permutation analyses only for WMC in N2pc time window ###

# Define N2pc time window 
n2pcTime <- c('start' = 220,
              'stop'  = 260)

# Compute mean amplitude within the time window
n2pcWmc <- trialSubtracted %>% 
  dplyr::group_by(wmc, id) %>% 
  dplyr::filter(time_point >= wmc.n2pc.time[['start']] & time_point <= wmc.n2pc.time[['stop']]) %>% 
  dplyr::summarise(signal = mean(subtracted))

# Perform permutation analyses through ez package (permutation version of t-test as we are comparing only two conditions)
ezPerm(data = n2pcWmc, 
       dv = signal, 
       wid = id, 
       between = wmc, 
       perms = 10000)

# Double cjheck with coin package
# coin::oneway_test(signal ~ wmc, data = n2pcWmc, distribution = approximate(10000))

# Plot
ezPlot(data = n2pcWmc, 
       dv = signal, 
       wid = id, 
       between = wmc.n2pc,
       x = wmc, 
       do_lines = FALSE, 
       do_bars = TRUE)
```

The results show that there is no significant difference in the N2pc time window between WMC groups. 

We then tested the differences in Ptc

```{r}
### Permutation analyses only for WMC in N2pc time window ###

# Define Ptc time window 
ptcTime <- c('start' = 280,
             'stop'  = 340)

# Compute mean amplitude within the time window
ptcWmc <- trial.subtracted %>% 
  dplyr::group_by(wmc, id) %>% 
  dplyr::filter(time_point >= wmc.ptc.time[['start']] & time_point <= wmc.ptc.time[['stop']]) %>% 
  dplyr::summarise(signal = mean(subtracted))

# Perform permutation analyses through ez package (permutation version of t-test as we are comparing only two conditions)
ezPerm(data = ptcWmc, 
       dv = signal, 
       wid = id, 
       between = wmc, 
       perms = 10000)

# Double cjheck with coin package
# coin::oneway_test(signal ~ wmc, data = ptcWmc, distribution = approximate(10000))

# Plot
ezPlot(data = ptcWmc, 
       dv = signal, 
       wid = id, 
       between = wmc.n2pc,
       x = wmc, 
       do_lines = FALSE, 
       do_bars = TRUE)
```

## Effects of Distractor Type

We observed that distractor circles elicited a late negativity. Through control of the scalp distributions (created in MATLAB), it appeared that the circle-driven negativity had similar neural generators with the L-driven N2pc. Thus, they could both be considered N2pc components. Consequently, we verified whether there was a difference in amplitude between the two. 

```{r}
### Compare L and O negativities

# Define time windows to compare (only for distractor effects)
timeWindows <- c('L_start' = 200,
                  'L_stop'  = 300,
                  'O_start' = 300,
                  'O_stop'  = 400)

# Extract mean amplitudes for the two negativities
negativitiesData <- trialSubtracted %>% 
  dplyr::filter(time_point >= timeWindows[['L_start']] & time_point <= timeWindows[['O_stop']]) %>% 
  dplyr::mutate(negativity = dplyr::case_when(time_point < timeWindows[['L_stop']] ~ 'L_negativity',
                                              time_point >= timeWindows[['O_start']] ~ 'O_negativity')) %>% 
  dplyr::group_by(id, distractor_type, negativity) %>% 
  dplyr::summarise(average = mean(subtracted)) %>% 
  dplyr::filter((distractor_type == 'L' & negativity == 'L_negativity') |
                (distractor_type == 'O' & negativity == 'O_negativity')) %>% 
  dplyr::ungroup()

### Compute analyses ###
# Set seed for replicability
set.seed(2021)
# Statistical analyses of the difference between the two negativities (should be null)
distrNegTest <- coin::oneway_test(average ~ distractor_type | id, 
                                    data         = negativitiesData, 
                                    distribution = approximate(10000))

# alternative
# ez::ezPerm(data = negativitieData, dv = average, wid = id, within = distractor_type, perms = 10000)
```

To fully understand the data, the activity elicited by circles in the time window 400ms-500ms was investigated as a possible late Ptc.

```{r}
### Late circle Ptc ###

# Extract mean amlitudes in defined window
latePtc <- trialSubtracted %>% 
  dplyr::filter((time_point >= 400 & time_point <= 500),
                distractor_type == 'O') %>% 
  dplyr::group_by(id) %>% 
  dplyr::summarise(avg = mean(subtracted)) %>% 
  tibble::add_column(time = 'ptc')

# Compute baseline value for comparison
baseline <- trialSubtracted %>% 
  dplyr::filter((time_point < 0),
                distractor_type == 'O') %>% 
  dplyr::group_by(id) %>% 
  dplyr::summarise(avg = mean(subtracted)) %>% 
  tibble::add_column(time = 'baseline')

late.ptc.all <- rbind(late.ptc, baseline)

# Compare late Ptc with baseline
coin::oneway_test(avg ~ factor(time) | id, data = late.ptc.all, distribution = approximate(10000))

# Alternatively
# ez::ezPerm(data = late.ptc.all, dv = avg, wid = id, within = time, perms = 10000)
```

