---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}
### Set up the environment  ###
# Load all the necessary packages
library(here)
library(tidyverse)
library(janitor)
library(ggplot2)
library(Rmisc)
library(RColorBrewer)
library(ggthemes)
library(permuco)
library(brms)
library(ggmcmc)
library(mcmc)
library(ggthemr)
# Plotting Theme
palette <- define_palette(
  swatch = c('#111111', '#E84646', '#233B43', '#C29365', '#168E7F', '#65ADC2', '#109B37', '#DB784D', '#f1c40f'),
  gradient = c(lower = '#E84646', upper = '#109B37'))
ggthemr(palette, layout = 'scientific')
# Distractors sides
right.distractors <- c(2,3,4,5,6,7,8)
left.distractors <- c(10,11,12,13,14,15,16)
```

```{r}
### LOAD AND PREPARE DATA ###
# Data was extracted through MATLAB and required some data wrangling
# HIGH WMC DATA
# Dataset should be contained in a folder called Datasets within the current directory
highWMCdata <- read_csv(here::here('Datasets', 'EEG_SingleTrial_High.csv')) %>% 
  # Pivot longer 
  tidyr::pivot_longer(cols      = '-200':dplyr::last_col(),
                      names_to  = 'time_point',
                      values_to = 'signal') %>% 
  # Clean column names (all lower case, style:name1_name2)
  janitor::clean_names() %>% 
  # Select relevant coloumns
  dplyr::select(trial,part_column, wmc_column, electrode, electrode_position, distractor_type, distractor_colour, distractor_position, time_point, signal) %>% 
  dplyr::rename(wmc = wmc_column,
                id  = part_column) %>% 
  dplyr::mutate(wmc = replace(wmc, wmc == 'hig', 'high'))
# LOW WMC DATA
# Dataset should be contained in a folder called Datasets within the current directory
lowWMCdata <- read_csv(here::here('SingleTrials_Datasets', 'Low_SingleTrial_All.csv')) %>% 
  # Pivot longer
  tidyr::pivot_longer(cols      = '-200':dplyr::last_col(),
                      names_to  = 'time_point',
                      values_to = 'signal') %>% 
  # Clean column names
  janitor::clean_names() %>% 
  dplyr::select(trial,part_column, wmc_column, electrode, electrode_position, distractor_type, distractor_colour, distractor_position, time_point, signal) %>% 
  dplyr::rename(wmc = wmc_column,
                id  = part_column) %>% 
  dplyr::mutate(wmc = replace(wmc, wmc == 'hig', 'high'))
# Join the data in a single dataset
trialData <- rbind(highWMCdata,
                   lowWMCdata) %>% 
  # Trasform relevant columns into factors
  dplyr::mutate(id                  = factor(id),
                wmc                 = factor(wmc),
                electrode           = factor(electrode),
                electrode_position  = factor(electrode_position),
                distractor_type     = factor(distractor_type),
                distractor_colour   = factor(distractor_colour),
                distractor_position = as_factor(distractor_position),
                time_point          = as.numeric(time_point))
```

# Compute subtraction

Ipsi - Contralateral in reference to the side of the presentation of the distractor
```{r}
### Compute subtracted waveforms ###
trialSubtracted <- trialData %>% 
  dplyr::select(!electrode) %>% 
  tidyr::pivot_wider(names_from  = electrode_position,
                     values_from = signal) %>% 
  # Compute contra - Ipsi
  dplyr::mutate(subtracted = contra - ipsi) %>% 
  dplyr::group_by()
```


# Bayesian Approach for N2pc
```{r}
### PREPARE DATA FOR MODEL ###

# Ptc Time Window
ptcTime <- c('start' = 280,
               'stop' = 340)

### PREPARE DATA ###
bayesData <- trialSubtracted %>% 
  dplyr::select(!c(ipsi, contra)) %>% 
  
  # Compute N2pc mean value by selecting only the relevant time window
  dplyr::filter(time_point >= ptcTime [['start']] & time_point <= ptcTime [['stop']]) %>% 
  dplyr::group_by(wmc, id, trial, distractor_type, distractor_colour) %>% 
  dplyr::summarise(n2pc_signal = mean(subtracted)) %>% 

  # Modify variable names for easy plotting and clean results
  dplyr::rename(type   = distractor_type,
                colour = distractor_colour) %>% 
  dplyr::mutate(wmc    = factor(wmc),
                type   = factor(type),
                colour = factor(colour),
                id     = factor(id))

# Set reference levels to Low WMC-Circle-Green
bayesData$wmc <- relevel(bayesData$wmc, ref = 'low')
bayesData$type <- relevel(bayesData$type, ref = 'O')
bayesData$colour <- relevel(bayesData$colour, ref = 'Gr')


```

```{r}
### DEFINE MODEL PARAMETERS ###
iterations       = 10000 # Number of iterations per chain
warmup           = 2500 # Number of burning iteration (length of warm up period that will be discarded)
ncores           = parallel::detectCores()# Number of RAM cores
chains           = 4    # Number of Markov Chains to create to estimate the prior distribution
seed             = 2021
inits            = 'random' # Randomly generates initial values
tapap            = .80 # Target-Average-Proposal-Acceptance-Probability (increase to decrease the step size during estimation)
```

The mixed effects Bayesian model is fitted using the package *brms* which interfaces R with STAN. 

```{r}
### CREATE MODEL ###
model1  <- brms::brm(formula = n2pc_signal ~ 1 + wmc * type * colour + (1|id),
                     data    = bayesData,
                     warmup  = warmup,
                     iter    = iterations,
                     chains  = chains,
                     inits   = inits,
                     cores   = ncores,
                     control = list(adapt_delta = tapap),
                     seed    = seed)
```

All the chains converged and no warnings were reported. The results of the model are:
```{r}
# Results of the model
summary(model1)
# Results using 95% HDI and ROPE
sjstats::equivalence_test(model1, ci = .95)
```

## Verify correct convergence of the model

In this section we will follow the When-to-Worry-and-How-to-Avoid-the-Misuse-of-Bayesian-Statistics – checklist (WAMBS-checklist) (Depaoli and van de Schoot, 2017) to verify that the model converged correctly and it is reliable.

1) The first step is to check that trace plots exhibit convergence. *brsm* did not output any warning message, but it is still preferable to visually verify this. 
```{r}
### TRACE PLOTS ###
model1.df <- ggmcmc::ggs(model1) # Transform model data into a long tibble
# Define parameters to plot
paramToPlot <- c('sigma',
                 'b_Intercept', 
                 'b_wmc', 
                 'b_typeL', 
                 'b_colourOr', 
                 'b_typeL:colourOr',
                 'b_wmchigh:typeL',
                 'b_wmchigh:colourOr',
                 'b_wmchigh:typeL:colourOr')

# Trace plots for main effects
trace1 <- model1.df %>% 
  dplyr::filter(Parameter %in% paramToPlot[1:5]) %>% 
  ggplot(aes(x   = Iteration,
             y   = value, 
             col = as.factor(Chain))) +
  geom_line(alpha = .5) +
  geom_vline(xintercept = warmup) +
  scale_colour_ggthemr_d() +
  facet_grid(Parameter ~ .,
             scale  = 'free_y',
             switch = 'y') +
  labs(title  = 'Chains Convergence',
       colour = 'Chains') +
  theme(strip.text.y.left = element_text(angle = 0))

# Trace plots for interactions
trace2 <- model1.df %>% 
  dplyr::filter(Parameter %in% paramToPlot[6:9]) %>% 
  ggplot(aes(x   = Iteration,
             y   = value, 
             col = as.factor(Chain))) +
  geom_line(alpha = .5) +
  geom_vline(xintercept = warmup) +
  scale_colour_ggthemr_d() +
  facet_grid(Parameter ~ .,
             scale  = 'free_y',
             switch = 'y') +
  labs(title = 'Chains Convergence',
       colour = 'Chains') +
  theme(strip.text.y.left = element_text(angle = 0))

# Plot trace plots together
cowplot::save_plot(paste(t, 'trace2.png'), plot = trace2)

# Use pp_check function
pp_check(model1, type = 'ecdf_overlay')
```

The trace plots show convergence much earlier than the end of the warm up period (vertical black line). We can also assess the Gelman-Rubin and Gewek's diagnostics to assess convergence. 

```{r}
model1.posterior <- as.mcmc(model1) # Extract info for diagnostic

# Gelman Diagnostic (Rhat)
coda::gelman.diag(model1.posterior[, 1:10])
coda::gelman.plot(model1.posterior[, 1:10])
```

The Gelman-Rubin's Diagnostic evaluates the within and between chain variability. As the upper C.I. are all 1, which is evidence for correct convergence. 

The Geweke's diagnostics computes a test for equality of means, in Z-scores, between the last and first portion of the chains. Thus, correct convergence should return a value less than |1.96|.

```{r}
coda::geweke.diag(model1.posterior[, 1:10])
coda::geweke.plot(model1.posterior[, 1:10])
```

The results support the above conclusions, further supporting correct convergence of the chains. 

2) The second step is to verify if the posterior distributions have enough information. It is possible to do so by inspecting the posterior distribution histograms. 

```{r}
# Posterior Distributions
stanplot(model1, type = 'hist')
```

The posterior distributions are centered and normally shaped with smooth slopes. Thus, the number of iterations was sufficient to extract enough information. 

3) The third step consists in evaluating the autocorrelation of the chains.

```{r}
coda::autocorr.diag(model1.posterior[, 1:10],
                    lags = c(0, 1,2,3,4,5,10,25,50))
```

4) Finally, posterior distribution can be visually analized to verify that they are reasonably meaningful. 

```{r}
### PLOT POSTERIOR DISTRIBUTIONS WITH ROPES AND HDI ###

# Extract HDI and ROPE
eqTest <- sjstats::equivalence_test(model1, ci = .95)
model1HDI <- eqTest[, c(1, 5, 7,8)] %>% 
  # Change model parameter names for nicer plotting
  dplyr::mutate(Parameter = case_when(Parameter == 'b_Intercept' ~ 'Intercept',
                                      Parameter == 'b_wmc' ~ 'WMC',
                                      Parameter == 'b_type' ~ 'Type',
                                      Parameter == 'b_colour' ~ 'Color',
                                      Parameter == 'b_wmc:type' ~ 'WMC:Type',
                                      Parameter == 'b_wmc:colour' ~ 'WMC:Color',
                                      Parameter == 'b_type:colour' ~ 'Type:Color',
                                      Parameter == 'b_wmc:type:colour' ~ 'WMC:Type:Color')) %>% 
  # Reorder factors for better plotting
  dplyr::mutate(Parameter = fct_relevel(Parameter, levels = c('WMC:Type:Color',
                                                              'Type:Color',
                                                              'WMC:Color',
                                                              'WMC:Type',
                                                              'Color',
                                                              'Type',
                                                              'WMC',
                                                              'Intercept')))


dataToPlot <- model1.df %>% 
  # Select factors of interest
  dplyr::filter(Parameter %in% c('b_Intercept', 
                                 'b_wmc', 
                                 'b_type', 
                                 'b_colour', 
                                 'b_wmc:type',
                                 'b_wmc:colour',
                                 'b_type:colour',
                                 'b_wmc:type:colour'),
                Iteration > warmup) %>% # Exclude warm up iterations as defined above (before model creation)
  # Standardize names as above
  dplyr::mutate(Parameter = case_when(Parameter == 'b_Intercept' ~ 'Intercept',
                                      Parameter == 'b_wmc' ~ 'WMC',
                                      Parameter == 'b_type' ~ 'Type',
                                      Parameter == 'b_colour' ~ 'Color',
                                      Parameter == 'b_wmc:type' ~ 'WMC:Type',
                                      Parameter == 'b_wmc:colour' ~ 'WMC:Color',
                                      Parameter == 'b_type:colour' ~ 'Type:Color',
                                      Parameter == 'b_wmc:type:colour' ~ 'WMC:Type:Color')) %>% 
  # Relevel factors as above
  dplyr::mutate(Parameter = fct_relevel(Parameter, levels = c('WMC:Type:Color',
                                                              'Type:Color',
                                                              'WMC:Color',
                                                              'WMC:Type',
                                                              'Color',
                                                              'Type',
                                                              'WMC',
                                                              'Intercept')))
bayesPlot1 <- dataToPlot %>%   
  ggplot() +
  ggridges::geom_density_ridges(aes(x      = value,
                                    y      = Parameter,
                                    height = ..density.., 
                                    fill   = Parameter,
                                    colour = Parameter),
                                scale          = 0.8, 
                                alpha          = 1) +
  geom_errorbarh(inherit.aes = FALSE,
                data         = model1HDI,
                aes(xmin     = HDI_low,
                    xmax     = HDI_high,
                    y        = Parameter,
                    colour   = Parameter), 
                size         = .6, 
                show.legend  = FALSE) +
  scale_colour_ggthemr_d() +
  annotate('rect',
           xmin    = -.61,
           xmax    = .61,
           ymin    = -Inf,
           ymax    = +Inf,
            fill   = 'red',
            colour = 'transparent',
            alpha  = .2) +
  geom_vline(xintercept = 0,
             colour     = 'red',
             size       = 1) +
  guides(colour = F) +
  labs(x = 'Ptc Mean Amplitude (mV)',
       fill = 'Predictors') +
  guides(fill = 'none')

```


