---
title: "LAI_Task_Behavioural"
author: "Daniele Scanzi"
date: "10/09/2021"
output: html_document
---

The following notebook contains the code used for analyses of Reaction Times and Accuracy in the LAI task (visual search task). Three predictors were included in the analyses:
* WMC: working memory capacity
* Distractor Type: L or Circle (O)
* Distractor Colour: Green (Gr) or Orange (Or)

```{r}
# Packages required for the analyses
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(psych)
library(pastecs)
library(PerformanceAnalytics)
library(cowplot)
library(car)
library(ggdist)
library(knitr)
library(coin)
library(ez)
library(ggthemr)
library(ggsignif)
library(parallel)

# Define theme for plotting using ggthemr
palette <- define_palette(
  swatch   = c('#111111', '#E84646', '#233B43', '#C29365', '#168E7F', '#65ADC2', '#109B37', '#DB784D'),
  gradient = c(lower = '#E84646', 
               upper = '#109B37'))
ggthemr(palette, layout = 'scientific')
```

```{r}
### Load Dataset and preapere it for analysis ###

# Load (data should be contained in a folder called Datasets within the current working directory)
originalData <- read_csv(here::here('Datasets', 'BehaviouralData.csv'))

# Define Working Memory Groups (extracted in previous analyses)
highWMC <- c(2, 3, 4, 9, 21, 27, 30, 31, 34)
lowWMC  <- c(6, 7, 8, 12, 14, 17, 20, 22, 23, 32, 37, 39, 41)

# Clean up the data. Eprime saves a bunch of variables that are useless and can be removed. 
cleanData <- copyData %>% 
  select('Subject',
         'Age',
         'Sex',
         'Handedness',
         'Procedure_Block_',
         'Trial',
         'Response_ACC',
         'Response_RT',
         'StimTrig',
         'StimTrig2',
         'newTargPos':'newDistance') %>% 
  
  # Standardize row names (all lower case; style word1_word2)
  janitor::clean_names() %>% 
  
  # Add WMC information for each participant
  dplyr::mutate(wmc_group = case_when(
    (subject %in% highWMC) ~ 'high',
    (subject %in% lowWMC)  ~ 'low'),
    new_distr_type = replace(new_distr_type, new_distr_type == '0', 'O' )) %>% # Some Os were saved as zeros
  # Select only LAI task trials
  dplyr::filter(procedure_block == 'BlockLAI') %>% 
  #Transform response times in numeric values
  dplyr::mutate(response_rt = as.integer(response_rt))

```

# Reaction Times Analyses
```{r}
### Extract mean reaction times for each condition ###
rtDataset <- cleanData %>% 
  # Group observations according to the three predictors for each subject 
  # WMC is not needed as each subject is associated to only one WMC group 
  # and we are not interested here in collapsing across groups (which requires to collapse across subjects)
  dplyr::group_by(subject, new_distr_type, new_distr_col) %>% 
  
  # Remove RT = 0, which probably are people that did not responded
  dplyr::filter(response_rt != 0) %>%
  
  #Compute mean for each subject
  dplyr::mutate(rt_mean = mean(response_rt)) %>%
  
  # Retain only useful columns
  dplyr::select(subject, wmc_group, new_distr_type, new_distr_col, rt_mean) %>% 
  
  # Remove duplicated rows 
  #dataset is in long-format, thus each subject has the same exact mean information in
  #each row in the new dataset
  distinct(subject, .keep_all = TRUE) %>% 
  dplyr::ungroup()
```

## RT Assumptions
```{r}
### Compute descriptives for RT based on WMC ###
rtDataset %>% 
  dplyr::group_by(wmc_group, new_distr_type, new_distr_col) %>% 
  dplyr::summarise(n       = n(),
            avg            = mean(rt_mean),
            median         = median(rt_mean),
            sd             = sd(rt_mean, na.rm = TRUE),
            se             = sd / sqrt(n), 
            min            = min(rt_mean),
            max            = max(rt_mean),
            IQR            = IQR(rt_mean), 
            skew           = skew(rt_mean),
            kurt           = kurtosi(rt_mean),
            shapiro.stat   = shapiro.test(rt_mean)$statistic,
            shapiro.pvalue = shapiro.test(rt_mean)$p.value)

### Assess QQ-Plots ###

# High WMC
rtDataset %>% 
  dplyr::filter(wmc_group == 'high') %>% 
  ggplot(aes(sample = rt_mean, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')

# Low WMC
rtDataset %>% 
  dplyr::filter(wmc_group == 'low') %>% 
  ggplot(aes(sample = rt_mean, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')


### Levene Test for Normality Assumption
car::leveneTest(rt_mean~wmc_group*new_distr_type*new_distr_col, 
           data = rtDataset)
```

## RT Permutation Analyses
```{r}
### Set up parameters for permutations ###
set.seed(2021) # For replicability
doParallel::registerDoParallel(cores = parallel::detectCores()) # Multi core computation for speed  (find and use avaiable cores)
nperm <- 10000 # Number of permutations
paral <- TRUE # Use parallel?
```

```{r}
### Perform Permutation Mixed-Design ANOVA type analysis ###
rtBehav <- ez::ezPerm(data      = rt.dataset, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = c(new_distr_type, new_distr_col), 
                       between  = wmc_group, 
                       perms    = nperm,
                       parallel = paral)
rtBehav
```
### Main Effects of Distractor Colour
```{r}
# RTs for distractor colour levels (pooling across all other dimensions)
rtDataset %>% 
  group_by(new_distr_col) %>% 
  summarise(n       = n(),
            rt.mean = mean(rt_mean),
            median  = median(rt_mean),
            sd      = sd(rt_mean, na.rm = TRUE), 
            se      = sd / sqrt(n), 
            min     = min(rt_mean),
            max     = max(rt_mean))

### Plot distributions of RT with means ad SE for distractor colour levels
rtDataset %>% 
  ggplot(aes(x      = new_distr_col, 
             y      = rt_mean,
             colour = new_distr_col)) +
  geom_jitter(width = 0.1, 
              alpha = 0.25) +
  stat_summary(fun  = 'mean', 
               geom = 'point',
               size = 3.2) +
  stat_summary(fun.data = 'mean_se',
               geom     = 'errorbar',
               width    = .2, 
               size     = 1) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)]) +
  # Add significance line
  geom_signif(comparisons = list(c('Gr', 'Or')), 
              annotations = '***',
              colour      = 'black') +
  labs(x = 'Distractor Color',
       y = 'Reaction Times (ms)') +
  scale_x_discrete(labels = c('Green', 'Orange')) +
  theme(legend.position = 'none')
```

### WMC:Distractor Type Interaction

This section includes the simple effects analyses for the significant interaction between WMC and distractor type.

```{r}
### Simple effects for the WMC on each level of Type ###

# Subset according to WMC group
rtWMC_high <- subset(rt.dataset, wmc_group == 'high') 
rtWMC_low  <- subset(rt.dataset, wmc_group == 'low')

# High WMC on each level of distractor
permRThigh <- ez::ezPerm(data   = rtWMC_high, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = new_distr_type,
                       perms    = nperm,
                       parallel = paral)

# High WMC on each level of distractor
permRTlow <- ez::ezPerm(data    = rtWMC_low, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = new_distr_type,
                       perms    = nperm,
                       parallel = paral)


### Simple effects for type on each level of WMC ###

# Subset according to WMC group
rtL <- subset(rt.dataset, new_distr_type == 'L')
rtO <- subset(rt.dataset, new_distr_type == 'O')

# High L on each level of WMC
permRTL <- ez::ezPerm(data    = rtL, 
                       dv       = rt_mean, 
                       wid      = subject,
                       between  = wmc_group,
                       perms    = nperm,
                       parallel = paral)

permRTO <- ez::ezPerm(data    = rtO, 
                       dv       = rt_mean, 
                       wid      = subject,
                       between  = wmc_group,
                       perms    = nperm,
                       parallel = paral)
```

```{r}
### WMC:Type interaction Plot ###
rtDataset %>% 
  ggplot(aes(x      = new_distr_type, 
             y      = rt_mean,
             group  = wmc_group,
             colour = wmc_group)) +
  geom_jitter(width = 0.1, 
              alpha = .25) +
  stat_summary(fun  = 'mean', 
               geom = 'point',
               size = 2.5) +
  stat_summary(fun  = 'mean',
               geom = 'line') +
  stat_summary(fun.data = 'mean_se',
               geom     = 'errorbar',
               width    = .08) +
  scale_colour_ggthemr_d() +
  # Add significance lines
  geom_signif(comparisons = list(c('L', 'O')), 
              annotations = '***',
              colour      = 'black') +
  annotate(geom    = 'text', 
           x       = 1.5,
           y       = 910, 
           label   = 'Both WMC groups',
              size = 2) +
  labs(x      = 'Distractor Type',
       y      = 'Reaction Times (ms)',
       colour = 'WMC') +
  scale_x_discrete(labels = c('L', 'Circle'))
```

```{r}
### Supplementary plot: type:colour ###
rtDataset %>% 
  ggplot(aes(x      = factor(new_distr_type), 
             y      = rt_mean,
             group  = new_distr_col,
             colour = new_distr_col)) +
  geom_jitter(width = 0.1, 
              alpha = 0.4) +
  stat_summary(fun  = 'mean', 
               geom = 'point',
               size = 3) +
  stat_summary(fun  = 'mean',
               geom = 'line') +
  stat_summary(fun.data = 'mean_se',
               geom     = 'errorbar',
               width    = .08) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      labels = c('Green', 'Orange')) +
  # Add significance lines
  geom_signif(comparisons = list(c('L', 'O')), 
              annotations = '***',
              colour      = 'black') +
  annotate(geom  = 'text', 
           x     = 1.5,
           y     = 1070, 
           label = 'Both colors',
           size  = 2) +
  # Manually add significance line
  geom_segment(mapping = aes(x = 0.85, y = 570, xend = 0.85, yend = 630), inherit.aes = F, size = 0.5) +
  geom_segment(mapping = aes(x = 0.85, y = 570, xend = 0.88, yend = 570), inherit.aes = F, size = 0.5) +
  geom_segment(mapping = aes(x = 0.85, y = 630, xend = 0.88, yend = 630), inherit.aes = F, size = 0.5) +
  geom_segment(mapping = aes(x = 2.15, y = 430, xend = 2.15, yend = 535), inherit.aes = F, size = 0.5) +
  geom_segment(mapping = aes(x = 2.12, y = 430, xend = 2.15, yend = 430), inherit.aes = F, size = 0.5) +
  geom_segment(mapping = aes(x = 2.12, y = 535, xend = 2.15, yend = 535), inherit.aes = F, size = 0.5) +
  # Add significance ***
  annotate(geom = 'text', x = 0.83, y = 600, label = '***', colour = 'black', angle = 90) +
  annotate(geom = 'text', x = 2.22, y = 480, label = '***', colour = 'black', angle = 90) +
  labs(x      = 'Distractor Type',
       y      = 'Reaction Times (ms)',
       colour = 'Colour') +
  scale_x_discrete(labels = c('L', 'Circle'))
```

# Accuracy Analyses

```{r}
### Extract accuracy values for each condition and participant ###
accDataset <- cleanData %>% 
  
  # Group observations according to the three predictors for each subject 
  # WMC is not needed as each subject is associated to only one WMC group 
  # and we are not interested here in collapsing across groups (which requires to collapse across subjects)
  group_by(subject, new_distr_type, new_distr_col) %>% 
  
  # Remove RT = 0, which probably are people that did not responded
  filter(response_rt != 0) %>%
  
  #Compute accuracy for each subject as number correct trials over tot of trials
  mutate(accuracy = sum(response_acc) / n()) %>% 
  # Retain only useful columns
  select(subject, wmc_group, new_distr_type, new_distr_col, accuracy) %>% 
  # Remove duplicated rows 
  # dataset is in long-format, thus each subject has the same exact mean information in
  # each row in the new dataset
  distinct(subject, .keep_all = TRUE)
```

## Accuracy Assumptions
```{r}
### Descriptives statistics for accuracy scores ###
accDataset %>% 
  group_by(wmc_group, new_distr_type, new_distr_col) %>% 
  summarise(n = n(),
            rt.mean        = mean(accuracy),
            median         = median(accuracy),
            sd             = sd(accuracy, na.rm = TRUE), 
            se             = sd / sqrt(n), 
            min            = min(accuracy),
            max            = max(accuracy),
            IQR            = IQR(accuracy), 
            skew           = skew(accuracy),
            kurt           = kurtosi(accuracy),
            shapiro.stat   = shapiro.test(accuracy)$statistic,
            shapiro.pvalue = shapiro.test(accuracy)$p.value)

### QQ-Plot ###

# High WMC
accDataset %>% 
  dplyr::filter(wmc_group == 'high') %>% 
  ggplot(aes(sample = accuracy, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')

# Low WMC
accDataset %>% 
  dplyr::filter(wmc_group == 'low') %>% 
  ggplot(aes(sample = accuracy, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')

### Levene Test for Normality Assumption
car::leveneTest(accuracy~wmc_group*new_distr_type*new_distr_col, 
           data = accDataset)
```

## Accuracy Permutation Analyses
```{r}
accBehav <- ez::ezPerm(data    = accDataset, 
                       dv       = accuracy, 
                       wid      = subject,
                       within   = c(new_distr_type, new_distr_col), 
                       between  = wmc_group, 
                       perms    = nperm,
                       parallel = paral)

accBehav
```

### Distractor Type Permutation Analyses
```{r}
### Descriptives for type levels (pooled across all other dimensions) ###
accDataset %>% 
  dplyr::group_by(new_distr_type) %>% 
  dplyr::summarise(n   = n(),
                   avg = mean(accuracy),
                   sd  = sd(accuracy),
                   SE  = sd / sqrt(n))
```

```{r}
### Plot distractor type distributions, means and SE ###
accDataset %>% 
  ggplot(aes(x      = new_distr_type, 
             y      = accuracy,
             colour = new_distr_type)) +
  geom_jitter(width = 0.1, 
              alpha = 0.4) +
  stat_summary(fun = 'mean', 
               geom = 'point',
               size = 3.2) +
  stat_summary(fun.data = 'mean_se',
               geom = 'errorbar',
               width = .08,
               size = 1) +
  scale_colour_ggthemr_d() +
  # Add significance lines
  geom_signif(comparisons = list(c('L', 'O')), 
              annotations = '***',
              colour      = 'black') +
  labs(x = 'Distractor Type',
       y = 'Reaction Times (ms)') +
  scale_x_discrete(labels = c('L', 'Circle')) +
  theme(legend.position = 'none')
```

