---
title: "Central_Task_Behavioural"
author: "Daniele Scanzi"
date: "15/10/2021"
output: html_document
---
The following notebook contains the code used for analyses of Reaction Times and Accuracy in the central task (control). Three predictors were included in the analyses:
* WMC: working memory capacity
* Distractor Type: L or Circle (O)
* Distractor Colour: Green (Gr) or Orange (Or)

```{r}
# Packages required for the analyses
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(psych)
library(pastecs)
library(PerformanceAnalytics)
library(cowplot)
library(car)
library(ggdist)
library(knitr)
library(coin)
library(ez)
library(ggthemr)
library(ggsignif)
library(parallel)

# Define theme for plotting using ggthemr
palette <- define_palette(
  swatch   = c('#111111', '#E84646', '#233B43', '#C29365', '#168E7F', '#65ADC2', '#109B37', '#DB784D'),
  gradient = c(lower = '#E84646', 
               upper = '#109B37'))
ggthemr(palette, layout = 'scientific')
```

```{r}
# Load
originalData <- read_csv(here::here('Datasets', 'BehaviouralData.csv'))

# Define Working Memory Groups (defined through previous analyses)
highWMC <- c(2, 3, 4, 9, 21, 27, 30, 31, 34)
lowWMC <- c(6, 7, 8, 12, 14, 17, 20, 22, 23, 32, 37, 39, 41)

# Clean up the data. Eprime saves a bunch of variables that are useless and can be removed. 
cleanData <- originalData %>% 
  select('Subject',
         'Age',
         'Sex',
         'Handedness',
         'Procedure_Block_',
         'Trial',
         'Response_ACC',
         'Response_RT',
         'StimTrig',
         'StimTrig2',
         'newTargPos':'newDistance') %>% 
  
  # Standardize row names (all lower case; style word1_word2)
  janitor::clean_names() %>% 
  
  # Add wmc information for each participant
  dplyr::mutate(wmc_group = case_when(
    (subject %in% highWMC) ~ 'high',
    (subject %in% lowWMC) ~ 'low'),
    new_distr_type = replace(new_distr_type, new_distr_type == '0', 'O' )) %>% # Some Os were saved as zeros
  
  # Select only centeal task trials
  dplyr::filter(procedure_block == 'BlockControl') %>% 
  
  #Trasform response times in numeric values
  dplyr::mutate(response_rt = as.integer(response_rt))
```

# Reaction Times Analyses
```{r}
rtDataset <- cleanData %>% 
  # Group observations according to the three predictors for each subject 
  # WMC is not needed as each subject is associated to only one WMC group 
  # and we are not interested here in collapsing across groups (which requires to collapse across subjects)
  dplyr::group_by(subject, new_distr_type, new_distr_col) %>% 
  
  # Remove RT = 0, which probably are people that did not responded
  dplyr::filter(response_rt != 0) %>%
  
  #Compute mean for each subject
  dplyr::mutate(rt_mean = mean(response_rt)) %>%
  # Retain only useful columns
  dplyr::select(subject, wmc_group, new_distr_type, new_distr_col, rt_mean) %>% 
  # Remove duplicated rows 
  # dataset is in long-format, thus each subject has the same exact mean information in
  # each row in the new dataset
  distinct(subject, .keep_all = TRUE)
```

## RT Descriptives
```{r}
### Compute descriptive statistics for each group ###
rtDataset %>% 
  dplyr::group_by(wmc_group, new_distr_type, new_distr_col) %>% 
  dplyr::summarise(n       = n(),
            avg            = mean(rt_mean),
            median         = median(rt_mean),
            sd             = sd(rt_mean, na.rm = TRUE), 
            se             = sd / sqrt(n), 
            min            = min(rt_mean),
            max            = max(rt_mean),
            IQR            = IQR(rt_mean), 
            skew           = skew(rt_mean),
            kurt           = kurtosi(rt_mean),
            shapiro.stat   = shapiro.test(rt_mean)$statistic,
            shapiro.pvalue = shapiro.test(rt_mean)$p.value)
```

## RT Permutation Analyses
```{r}
### Set up parameters for permutations ###
set.seed(2021) # For replicability
doParallel::registerDoParallel(cores = parallel::detectCores()) # Multi core computation for speed  (find and use avaiable cores)
nperm <- 10000 # Number of permutations
paral <- TRUE # Use parallel?
```

```{r}
### Perform Permutation Mixed-Design ANOVA type analysis ###
rtBehav <- ez::ezPerm(data      = rtDataset, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = c(new_distr_type, new_distr_col), 
                       between  = wmc_group, 
                       perms    = nperm,
                       parallel = paral)
rtBehav
```

### Main effects of colour
```{r}
### Plot distribution, mean and SE for levels of distractor colour
rtDataset %>% 
  ggplot(aes(x      = new_distr_col,
             y      = rt_mean,
             colour = new_distr_col)) +
  geom_jitter(width = 0.1, 
              alpha = 0.25) +
  stat_summary(fun = 'mean', 
               geom = 'point',
               size = 3.2) +
  stat_summary(fun.data = 'mean_se',
               geom     = 'errorbar',
               width    = .2, 
               size     = 1) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)]) +
  # Add significance lines
  geom_signif(comparisons = list(c('Gr', 'Or')), 
              annotations = '***',
              colour      = 'black') +
  labs(x = 'Distractor Color',
       y = 'Reaction Times (ms)') +
  scale_x_discrete(labels = c('Green', 'Orange')) +
  theme(legend.position = 'none')
```

### WMC and Distractor type Interaction

This section includes the simple effects analyses for the significant interaction between WMC and distractor type.

```{r}
### Simple effects for the WMC on each level of Type ###

# Subset according to WMC group
rtWMC_high <- subset(rt.dataset, wmc_group == 'high') 
rtWMC_low  <- subset(rt.dataset, wmc_group == 'low')

# High WMC on each level of distractor
permRThigh <- ez::ezPerm(data   = rtWMC_high, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = new_distr_type,
                       perms    = nperm,
                       parallel = paral)

# High WMC on each level of distractor
permRTlow <- ez::ezPerm(data    = rtWMC_low, 
                       dv       = rt_mean, 
                       wid      = subject,
                       within   = new_distr_type,
                       perms    = nperm,
                       parallel = paral)


### Simple effects for type on each level of WMC ###

# Subset according to WMC group
rtL <- subset(rt.dataset, new_distr_type == 'L')
rtO <- subset(rt.dataset, new_distr_type == 'O')

# High L on each level of WMC
permRTL <- ez::ezPerm(data    = rtL, 
                       dv       = rt_mean, 
                       wid      = subject,
                       between  = wmc_group,
                       perms    = nperm,
                       parallel = paral)

permRTO <- ez::ezPerm(data    = rtO, 
                       dv       = rt_mean, 
                       wid      = subject,
                       between  = wmc_group,
                       perms    = nperm,
                       parallel = paral)
```

```{r}
### WMC:Type interaction Plot ###
rtDataset %>% 
  ggplot(aes(x      = new_distr_type, 
             y      = rt_mean,
             group  = wmc_group,
             colour = wmc_group)) +
  geom_jitter(width = 0.1, 
              alpha = .25) +
  stat_summary(fun = 'mean', 
               geom = 'point',
               size = 2.5) +
  stat_summary(fun  = 'mean',
               geom = 'line') +
  stat_summary(fun.data = 'mean_se',
               geom     = 'errorbar',
               width    = .08) +
  scale_colour_ggthemr_d() +
  # Add significance lines
  geom_signif(comparisons = list(c('L', 'O')), 
              annotations = '***',
              colour      = 'black') +
  annotate(geom    = 'text', 
           x       = 1.5,
           y       = 1600, 
           label   = 'Both WMC groups',
              size = 2) +
  labs(x = 'Distractor Type',
       y = 'Reaction Times (ms)',
       colour = 'WMC') +
  scale_x_discrete(labels = c('L', 'Circle'))
```

# Accuracy Analyses
```{r}
### Extract accuracy values for each condition and participant ###
accDataset <- cleanData %>% 
  
  # Group observations according to the three predictors for each subject 
  # WMC is not needed as each subject is associated to only one WMC group 
  # and we are not interested here in collapsing across groups (which requires to collapse across subjects)
  group_by(subject, new_distr_type, new_distr_col) %>% 
  
  # Remove RT = 0, which probably are people that did not responded
  filter(response_rt != 0) %>%
  
  #Compute accuracy for each subject as number correct trials over tot of trials
  mutate(accuracy = sum(response_acc) / n()) %>% 
  # Retain only useful columns
  select(subject, wmc_group, new_distr_type, new_distr_col, accuracy) %>% 
  # Remove duplicated rows 
  # dataset is in long-format, thus each subject has the same exact mean information in
  # each row in the new dataset
  distinct(subject, .keep_all = TRUE)
```

## Accuracy Descriptives
```{r}
### Descriptives statistics for accuracy scores ###
accDataset %>% 
  group_by(wmc_group, new_distr_type, new_distr_col) %>% 
  summarise(n = n(),
            rt.mean        = mean(accuracy),
            median         = median(accuracy),
            sd             = sd(accuracy, na.rm = TRUE), 
            se             = sd / sqrt(n), 
            min            = min(accuracy),
            max            = max(accuracy),
            IQR            = IQR(accuracy), 
            skew           = skew(accuracy),
            kurt           = kurtosi(accuracy),
            shapiro.stat   = shapiro.test(accuracy)$statistic,
            shapiro.pvalue = shapiro.test(accuracy)$p.value)

### QQ-Plot ###

# High WMC
accDataset %>% 
  dplyr::filter(wmc_group == 'high') %>% 
  ggplot(aes(sample = accuracy, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')

# Low WMC
accDataset %>% 
  dplyr::filter(wmc_group == 'low') %>% 
  ggplot(aes(sample = accuracy, 
             colour = new_distr_col,
             fill   = new_distr_col,
             group  = new_distr_col)) +
  stat_qq(show.legend = FALSE) +
  stat_qq_line(size  = .8,
               alpha = .8) +
  scale_colour_manual(values = palette[['swatch']][c(7,8)],
                      name   = 'Distractor Colour') +
  facet_wrap(~new_distr_type) +
  labs(y = 'Sample Quantile')

### Levene Test for Normality Assumption
car::leveneTest(accuracy~wmc_group*new_distr_type*new_distr_col, 
           data = accDataset)
```

## Accuracy Permutation Analyses

Note: the accuracy scores appeared to be at chance level for each group. Thus, the analyses showed no significant effects. 

```{r}
accBehav <- ez::ezPerm(data     = accDataset, 
                       dv       = accuracy, 
                       wid      = subject,
                       within   = c(new_distr_type, new_distr_col), 
                       between  = wmc_group, 
                       perms    = 10000,
                       parallel = TRUE)

accBehav
```

